---
title: "pa3"
author: "Kaylee Fernandez"
date: "2025-02-28"
output: html_document
---

# load packages

```{r}
library("tidyverse")
library("untidydata")
library("ggplot2")

data(package = "untidydata") 
data(spanish_vowels, package = "untidydata") 
```

# explore

```{r}
str(spanish_vowels)
summary(spanish_vowels)
glimpse(spanish_vowels)

```

# tidying the data

```{r}
tidy_spanish_vowels <- spanish_vowels |>
  separate(
    col = label,
    into = c("ID", "gender", "vowel"),
    sep = "-" 
  )

print(tidy_spanish_vowels)
```

# descriptive statistics

```{r}

tidy_spanish_vowels |>
  group_by(gender, vowel) |> #grouping by participant type and item
  summarize(
    avg_f1 = mean(f1),
    avg_f2 = mean(f2),
    sd_f1 = sd(f1),
    sd_f2 = sd(f2), 
    ) |>
  print(n = Inf)
```

# linear model

```{r}
model_spanish_vowels <- lm(formula = f1 ~ f2 : vowel, data = tidy_spanish_vowels)
summary(model_spanish_vowels)
```

# plot including regression line

```{r}
tidy_spanish_vowels |>
  ggplot() + 
  aes(x =f2, y = f1, color = vowel) + # mapping the x and y axis
  geom_point() + # adding the geometric object
  geom_smooth(method = "lm") + # adding the regression line
  labs(title = "F1 as a Function of F2 by vowel")
```

## general observations

The dataset includes 25 male and 25 female Spanish speakers, each producing the vowels /a, e, i, o, u/ three times. F1 and F2 measurements were recorded for each vowel. On average, females had higher F1 and F2 values than males. A linear model was fitted to examine the interaction between F2 and vowel (predictors) on F1 (dependent variable). The results revealed significant interactions between F2 and each vowel on F1 (p < .001). The regression plot shows distinct vowel clusters with positive slopes, indicating that for each vowel, higher F2 values are associated with higher F1 values.
